The title

### 2016-10-10 analysis/paper/paper.Rmd
new paragraph and sub-headings

Since the 19th century, the focus of archaeological interest lay on the central part of the Irrawaddy River in central Myanmar. The first artefacts were discovered by Morris in 1930 in Upper Myanmar and later assigned to the “new” Anyathian culture from the Lower Palaeolithic in Southeast Asia. The Anyathian artefacts are mainly not found in situ but in the secondary deposits of the Pleistocene terrace gravels of the Irrawaddy River. The occurrence of only few handaxes is one main characteristic of the Anyanthian culture. The occupation of the Irrawaddy River terraces during the Pleistocene seems to be limited to the east bank of the River and the central part of Myanmar because the finds concentrate in the region around Chauk, Nyaung-U and Yenanyaung [@movius1944early; @deTerra1943].

The three phases of the Early Anyathian are all dominated by a high degree of uniformity of the stone tools, probably due to raw material constraints (fossil wood and silicified tuff). Fossil wood is difficult to flake because it easily flakes parallel to the natural fibres of the wood. Most of the artefacts made from fossil wood are worked on only one plane; multiple flake scars are rare. Silicified tuff is more suitable for stone tool production due to its fine-grained homogenous texture and conchodial fractures, although it can be very porous. In total, 483 implements were found (261 from fossil wood, 220 from silicified tuff and 2 from quartzite). Implements made from fossil wood display a limited range of tools, including mostly handaxes, as well as choppers and chopping tools. Generally, stone tools made of silicified tuff are more massive than tools made from fossil wood. Choppers are dominant, whereas there are few handaxes [@deTerra1943].

In addition, 23 early Anyathian artefacts were found in situ in the lateritic gravel and gravel deposits of terrace 1 near Chauk on the Hill of Chinaungma. Movius (1943) assigned these implements to the first phase of the Early Anyathian and the beginning of the Middle Pleistocene as the oldest human artefacts found in the Irrawaddy Valley. Implements of phase 2 of the Early Anyathian show either a ferruginous crust, or are heavily rolled. The former implies an occupation of this area during the forming of the crust in a long interpluvial period. The latter suggests that these implements were derived from older deposits. Most of the Artefacts from the third phase of the Early Anyathian are very heavily rolled and are associated with the early Upper Pleistocene when the gravels were deposited [@deTerra1943].

Norton et al. [-@norton2006middle] reworked the concept of the Movius Line by incorporating existing issues with the “traditional” Movius Line. The three main characteristics of the “Movius Line sensu lato” are the lower frequency of handaxe bearing sites in East Asia compared to Africa and India; the much lower percentage of bifacially made tools in East Asian assemblages and the morphological similarities to Acheulean artefacts [@norton2006middle; @norton2008movius].

>>>>

### First archaeological exploration and the Movius Line

Since the 19th century, archaeological interest was focussed on the central part of the Irrawaddy River in central Myanmar. The first artefacts were discovered by Morris in 1930 in Upper Myanmar and later assigned to the “new” Anyathian culture from the Lower Palaeolithic in Southeast Asia. The Anyathian artefacts are mainly in the secondary deposits of the Pleistocene terrace gravels of the Irrawaddy River. The occurrence of only few handaxes is a main characteristic of Anyanthian assemblages. The occupation of the Irrawaddy River terraces during the Pleistocene seems to be limited to the east bank of the River and the central part of Myanmar because the finds concentrate in the region around Chauk, Nyaung-U and Yenanyaung [@movius1944early; @deTerra1943].

Due to raw material constraints (fossil wood and silicified tuff), the three phases of the Early Anyathian are dominated by a high degree of uniformity of the stone tools. Fossil wood is difficult to flake because it easily flakes parallel to the natural fibres of the wood. Most of the artefacts made from fossil wood are worked on only one plane; multiple flake scars are rare. Silicified tuff is more suitable for stone tool production due to its fine-grained homogenous texture and conchodial fractures, although it can be very porous. In total, 483 implements were found (261 from fossil wood, 220 from silicified tuff and 2 from quartzite). Implements made from fossil wood include mostly handaxes, as well as choppers and chopping tools. Generally, stone tools made of silicified tuff are more massive compared to fossil wood. Choppers are dominant, whereas there are few handaxes [@deTerra1943].

In addition, 23 Early Anyathian artefacts were found in situ in the lateritic gravel and gravel deposits of terrace 1 near Chauk on the Hill of Chinaungma. Movius (1943) assigned these implements to the first phase of the Early Anyathian and the beginning of the Middle Pleistocene as the oldest human artefacts found in the Irrawaddy Valley. Implements of phase 2 of the Early Anyathian show either a ferruginous crust, or are heavily rolled. The former implies an occupation of this area during the forming of the crust in a long interpluvial period. The latter suggests that these implements were derived from older deposits. Most of the Artefacts from the third phase of the Early Anyathian are very heavily rolled and are associated with the early Upper Pleistocene when the gravels were deposited [@deTerra1943].

Compared to the more complex technologies found in Africa, Europe, the Levant and western/southern Asia and the simpler chopper/chopping tool technology of Southeast/East Asia, Movius proposed the concept of the Movius Line separating the Lower Palaeolithic into two cultures - the crude or non- stone tool-making cultures in E and SE Asia (e.g. Imjin/Hantan River Basin, Korea and Luonan Basin and Longyadong Cave, China) and northern China (e.g. Zhoukoudian) and the more sophisticated stone tool-making cultures in the Old World (e.g. Olduvai George Beds, Tanzania and Olorgesailie, Kenya) [@brumm2012biface2; @norton2006middle].

### The Movius Line in modern quaternary science

In modern quaternary science, the Movius Line is highly debated. Norton et al. [-@norton2006middle] reworked the concept of the Movius Line by incorporating existing issues with the “traditional” Movius Line. The three main characteristics of the “Movius Line sensu lato” are the lower frequency of handaxe bearing sites in East Asia compared to Africa and India; the much lower percentage of bifacially made tools in East Asian assemblages and the morphological similarities to Acheulean artefacts [@norton2006middle; @norton2008movius].
